---
title: "Practical 1"
format: html
editor: visual
---

## Setup/Preamble

```{r}
#| warning: false
library(tidyverse)
library(palmerpenguins)
library(nycflights13)
```

## Section 1

Question 1

```{r}
glimpse(penguins)
```

Rows: 344 Columns: 8

Question 2

```{r}
ggplot(penguins, aes(x = bill_length_mm, y = bill_depth_mm)) +
  geom_point()
```

Question 3

```{r}
ggplot(penguins, aes(x = bill_length_mm, y = bill_depth_mm)) +   
  geom_point() +
  labs(
    caption = "Data come from the palmerpenguins package."
  )
```

Question 4

```{r}
ggplot(penguins, aes(x = species, y = bill_depth_mm)) +   
  geom_point()
```

A better geom to use would be a boxplot.

Question 5

```{r}
ggplot(penguins, aes(x = species, y = bill_depth_mm)) +   
  geom_boxplot()
```

Question 6

```{r}
ggplot(penguins, aes(x = flipper_length_mm, y = body_mass_g)) +   
  geom_point(aes(color = bill_depth_mm), size = 3) +
  geom_smooth() # method = 'loess' by default
```

## Section 2

Question 7

It is a good idea to understand the range of values of `carat` before choosing bin sizes.

```{r}
diamonds |>
  summary() # for the whole data frame

summary(diamonds$carat) # for just the single column
```

Given that `carat` ranges from 0.20 to 5.01, lets use 0.01, 0.10 and 1 as bin widths.

```{r}
ggplot(diamonds, aes(x = carat)) +   
  geom_histogram(binwidth = 0.1) # adjust bin width and re-run.
```

## Section 3

Question 8

```{r}
ggplot(mpg, aes(x = displ, y = hwy)) +      
  geom_point()
```

Numerical variable chosen is `cty` (city mileage).

```{r}
ggplot(mpg, aes(x = displ, y = hwy, color = cty)) +      
  geom_point()
```

```{r}
ggplot(mpg, aes(x = displ, y = hwy, size = cty)) +      
  geom_point()
```

```{r}
ggplot(mpg, aes(x = displ, y = hwy, color = cty, size = cty)) +      
  geom_point()
```

```{r}
ggplot(mpg, aes(x = displ, y = hwy, color = cty, size = cty, shape = drv)) +   
  geom_point()
```

The difference is a numerical variable doesn’t work with the shape aesthetic but a categorical variable does.

## Section 4

Question 9

```{r}
# Had an arrival delay of two or more hours
flights |>
  filter(arr_delay >= 120) |>
  arrange(desc(arr_delay))

# Went to Houston (IAH or HOU)
flights |>
  filter(dest %in% c("IAH", "HOU"))

# Were operated by United, American, or Delta
flights |>
  filter(carrier %in% c("UA", "DL", "AA"))

# Departed in July, August, or September
flights |>
  filter(month %in% c(7, 8, 9))

# Arrived 2+ hours late, but left on time
flights |> 
  filter(arr_delay >= 120 & dep_delay <= 0)

# Were delayed by at least 1 hour, but gained over 30 mins in air
flights |> 
  filter(dep_delay >= 60 & dep_delay - arr_delay > 30)
```

Question 10 Yes, there was a flight on every day of 2013 since there are 365 distinct combinations of year, month, and day, which is equal to the number of days in the year 2013.

```{r}
flights |> 
  distinct(year, month, day) |>
  nrow()
```

Question 11

```{r}
flights |> 
  arrange(desc(distance)) |> # arranging distance largest to smallest
  relocate(distance) # optional line
```